Multimodal Language Learning for Object Retrieval in the Face of Missing Modalities

### 2023-02-08 Paper/main-SIGIR.tex
Paper submitted to SIGIR 23

\documentclass[10pt]{article} % For LaTeX2e
\usepackage{tmlr}
%%\usepackage{showframe}
% If accepted, instead use the following line for the camera-ready submission:
% \usepackage[accepted]{tmlr}
% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
% \usepackage[preprint]{tmlr}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}

% My Packages
\usepackage{cleveref}
\usepackage{comment}
\usepackage{csvsimple}
\usepackage{datatool}
\usepackage{todonotes}[disable]
\usepackage{graphicx}
\usepackage{subcaption}
\newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
\newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
% \newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
\newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
\newcommand{\todocmi}[1]{\todo[inline,color=green!40]{\small #1 -- Cynthia}}
\newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
\newcommand{\todoffinline}[1]{\todo[inline,color=blue!20]{\small #1 -- Frank}}
\newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}

@@

\title{Multimodal Language Learning in the Face of Missing Modalities}
% \\ Extended Multimodal Alignment \\ Combining Geometric and classification ... }

% Authors must not appear in the submitted version. They should be hidden
% as long as the tmlr package is used without the [accepted] or [preprint] options.
% Non-anonymous submissions will be rejected without review.

\author{\name Kasra Darvish \email kasradarvish@umbc.edu \\
      \addr Department of Computer Science\\
      University of Maryland Baltimore County
      \AND
      \name Edward Raff \email Raff\_Edward@bah.com \\
      \addr University of Maryland Baltimore County \\
      Booz Allen Hamilton
      \AND
      \name Francis Ferraro \email ferraro@umbc.edu \\
      \addr University of Maryland Baltimore County
      \AND
      \name Cynthia Matuszek \email cmat@umbc.edu\\
      \addr University of Maryland Baltimore County \\
      CIFAR Fellow}

% The \author macro works with any number of authors. Use \AND 
% to separate the names and addresses of multiple authors.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\def\month{MM}  % Insert correct month for camera-ready version
\def\year{YYYY} % Insert correct year for camera-ready version
\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version

@@

\maketitle

\begin{abstract}
    % Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval. Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment. However, physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks. In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities. We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command. We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.

    % shorter version

    Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities with which a human may interact, both to communicate a desired goal and for the agent to recognize a desired target object. For such problems, there has been little research on integrating more than two modalities. While there have been widely popular works based on cross-entropy, there is an entirely separate family of approaches based on explicit geometric alignment known as contrastive learning. However, to the best of our knowledge there has been no work on combining the two approaches for multimodal learning. We propose to combine both families of approaches and argue that the two are complementary. We introduce \textit{extended multimodal alignment (EMMA)}, a geometric (distance-based) method combined with a contrastive loss function. Our method can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable. We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command. We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model converges approximately 5 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning. % (in 80\% less time).
    %when modalities are ablated. 
    %TODO: another baseline where we use cross-entropy loss only is missing I guess.
    The code is publicly available on GitHub and will be included for the camera-ready version (it is redacted for anonymity). 
\end{abstract}

\section{TODO}
\begin{itemize}

    \item \TD modify it for SIGIR
    \item \TD EMMA as a contstrained optimization for Geometric portion to not satisfy it.
    \item \TD constrained SupCon per instance modality constraints: for this particular anchor and corresponding neighbors, we want this constraint to be satisfied.
    \item \OK Do you control for in batch negatives for supcon portion? Yes.
    \item \TD alternate training of supcon and emma. E.g. train supcon for 50 epochs and emma for the next 50 epochs
    \item \TD removing objects that have less than a number of instances? yes, but it's not a good story, since we want emma to work in low-data regime.
    \item \TD ablating number of data points and have a graph based on different number of training data --> maybe EMMA degrades less than SupCon and that would be good
    \item \TD use supcon portion to reweight the geometric values or use a method to have different scores for different pairs in the geometric loss. Targeting specific pair of embeddings.
    \item \TD it seems to be a tension between geometric portion and the supcon loss
    \item \OK another column for number of NAs in test and train and then correlation between them

    \item \OK add another column for the number of instances and then compute correlation between that and the differential scores. Take the correlation between number of instances and each column (metric) and then the correlation and the average of columns.

    \item listen to the speeches for NA descriptions and the ones where there is a significant difference in performance

    \item \OK listen to the speeches for empty transcriptions.

    \item \TD take a look at the depth images

    \item going forward we have two options: Improve the model, or do a deep analysis why our results are like this

    \item We could do something that addresses the "open questions" (which questions remain?)

    I guess the main open question is the mixed results and the fact that they are within the std dev of each other. Basically, I think they are unsure whether being faster is a good result or not.
    What the reader would take away?

    \item \DO What the rich analysis would consist of?

    \item Why are we getting what we are getting?

    Visualize individual data points and what are the failure modes.
    You can do the same for type by type (apples vs books)
    The percent that EMMA got right and SupCon got wrong and vice versa.

    have pictures, positive set and negative set,
    and all modalities and then show scores for all of them.

    \item \TD Different datasets or tasks
    \item \TD How can we tell when modalities are complementary?

    I think they meant when loss functions are complementary, am I wrong?

    \item \TD How can we tell what losses to use in what cases?

    \item  \TD Why didn't our approach work better, and what can we do to improve it?

    One reason could be that \supcon{} considers the distance between all pairs of data points in a batch.

    \item What's the next thing to try or add?

    We could add a weighting term to balance the effect of \supcon{} and \geom{}.

\end{itemize}

\section{Introduction}
\label{intro}

Inspired by the multimodal nature of human interaction with the world, it is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources. The benefits of integrating multiple modalities are twofold; first, complementary information can be extracted from different modalities that can help with understanding the world, and second, additional modalities can help in the cases when one or more sources of data about the world become unavailable.
% 
Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval. Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\citep{hong2021gilbert,Zhuge_2021_CVPR_VLP}.

Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB plus depth images) with single labels, such as those provided by textual natural language. Simultaneously using additional inputs from different modalities is an underexplored area, in part due to the domain-specific nature of such \textit{n}-ary learning approaches. With the modern proliferation of audio and text based communication and home agents (e.g., Alexa/Google Home), there is a growing need to handle more modalities, and simultaneously their potential failures.

One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information. Hardware can become damaged or defective, sensors can get blocked or obstructed, and various adverse but not uncommon conditions can remove a modality from use.
Current multimodal approaches are typically not robust to the loss of one or more modalities at test time, as may happen if, for example, a physical agent fails to retrieve data from a particular sensor. In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.

We consider the domain of grounded language-based object retrieval~\citep{hu2016natural, triplet_loss_2021_CVPR}, in which objects in an environment must be identified based on linguistic instructions. This can be considered a special case of image retrieval~\citep{huang2017deep, ma2020large, novak2015large, vo2019composing} in which objects are identified using visual inputs in combination with other sensor modalities. Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or  speech~\citep{RichardsDarvishMatuszekCategoryFree20}. 
However, despite the multisensory nature of object retrieval, much of the existing work has not previously been extended to include an arbitrary number of modalities.

% \todo[inline]{Whats the name of our method? When I skim the paper its hard to tell what ``the new thing is'', and then I have to read. I think being explicit here, with something like ``\textit{we propose our new method BLA}'' will help -Ed}

The ultimate aim of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities. In this work, we propose our new method called \textit{extended multimodal alignment (\ours{})}, a generalized geometric method combined with a cross-entropy based supervised contrastive loss function~\cite{NEURIPS2020_supervised_contrastive}, which we abbreviate \supcon{}. Our method can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data. We demonstrate steps towards this ultimate goal with a mechanism that learns quickly from a variety of input signals and is robust to ablation of certain inputs. This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.

Our contributions are as follows:
\begin{enumerate}
    \item Proposing a new approach to multimodal learning by combining geometric and cross-entropy methods and showing that it outperforms and converges faster compared to each method separately.
    % \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.

    \item Our method is designed to be applicable to additional modalities as they become available.

    \item Our method converges approximately 5 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning~\citep{chen2020simple} and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} models when all modalities are available.

    \item Demonstrating robustness in the face of missing input signals when one or more modalities are ablated at test time, and proposing a simple averaging method to quantify performance for multimodal object retrieval scenarios.

    \item Demonstrating the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.

    % \item Stable model that outperforms state-of-the-art contrastive learning models when batch size decreases.\todoff{add motivation in the intro about why robustness to a smaller batch size is important}

\end{enumerate}

@@

There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case. In this task, language is used to formulate queries against datasets of images, for example in text-and-image matching tasks for fashion data~\citep{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\citep{huang2017deep}, and general photographs of objects~\citep{ma2020large, novak2015large, hong2021gilbert}. Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\citep{vo2019composing}. While recent works have focused on grounding with respect to a knowledge set \citep{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning to perform grounding over an arbitrary number of input modalities.

The growing number of datasets that contain modalities beyond images and text demonstrates the importance and the challenge of this task. \citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition. They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\citep{KebeAAAI2022}. 
\citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization. They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}. In this work, we focus on alignment, the process of identifying the direct relationship between elements across modalities.

We highlight a key difference with recent work like \citet{10.1145/3397271.3401232} and \citet{10.1145/3331184.3331213} who develop multi-modal retrieval models. These retrieval models are class based, where the model is trained to recognize any object of the same retrieval class as being the `same.' Our method and data is instance based, where different objects are not considered the same even if they have the same class, due to the grounding goals of our work (i.e., the agent should identify the specified object, not equivalent objects). These works also do not consider performance in the case of failure of an input modality.

A number of different approaches have been proposed for linguistic-based alignment learning. \citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language. Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video, and these embeddings are then mapped to a coarse space where text is also mapped. Their network respects specificity and comparability of different modalities. The loss function they define consists of two distance-based terms for the corresponding spaces, differing from ours in that they do not handle arbitrarily many modalities and do not focus on robustness in the face of modality drop-outs. \citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks. \citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss \citep{Chechik:2010:LSO:1756006.1756042}, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.

In general there are two different families of approaches to multimodal learning in the literature. One family is based on classification and cross-entropy losses such as~\citet{NEURIPS2020_supervised_contrastive, chen2020simple} and the other family is geometric based such as~\citet{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}.
In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy based \supcon{} loss function is superior.

% \todokdinline{Should we move the following two paragraphs to the Approach section?}
% \todocmi{Actually yes, although I think it's more like 3 paragraphs. Rewriting the subsequent paragraphs so they don't appear out of nowhere would help.}
% \todokdinline{After moving baselines to the approach section and rewriting it, I think it's better if we leave these here. Let me know if you think otherwise.}

>>>>

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.
% \documentclass[sigconf,review,natbib=true]{acmart}
\documentclass[sigconf,natbib=true,anonymous=true]{acmart}
%% NOTE that a single column version may be required for 
%% submission and peer review. This can be done by changing
%% the \doucmentclass[...]{acmart} in this template to 
%% \documentclass[manuscript,screen]{acmart}
%% 
%% To ensure 100% compatibility, please check the white list of
%% approved LaTeX packages to be used with the Master Article Template at
%% https://www.acm.org/publications/taps/whitelist-of-latex-packages 
%% before creating your document. The white list page provides 
%% information on how to submit additional LaTeX packages for 
%% review and adoption.
%% Fonts used in the template cannot be substituted; margin 
%% adjustments are not allowed.

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[SIGIR '23]{}{2023}{Taipei, Taiwan}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
% My Packages
\usepackage{cleveref}
\usepackage{comment}
\usepackage{csvsimple}
\usepackage{datatool}
\usepackage{todonotes}[disable]
\usepackage{graphicx}
\usepackage{subcaption}
\newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
\newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
% \newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
\newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
\newcommand{\todocmi}[1]{\todo[inline,color=green!40]{\small #1 -- Cynthia}}
\newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
\newcommand{\todoffinline}[1]{\todo[inline,color=blue!20]{\small #1 -- Frank}}
\newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}

@@

% Remove for submission!!
% Remove for submission!!
% Remove for submission!!
% Remove for submission!!
\pagenumbering{arabic} 
\settopmatter{printfolios=true}

@@

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Multimodal Language Learning for Object Retrieval in the Face of Missing Modalities}
% \\ Extended Multimodal Alignment \\ Combining Geometric and classification ... }

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Kasra Darvish}
% \authornote{Both authors contributed equally to this research.}
% \email{kasradarvish@umbc.edu}
% \orcid{1234-5678-9012}
% \author{G.K.M. Tobin}
% \authornotemark[1]
% \email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{University of Maryland Baltimore County}
  % \streetaddress{P.O. Box 1212}
  \city{Baltimore}
  \state{MD}
  \country{USA}
  % \postcode{43017-6221}
}
\email{kasradarvish@umbc.edu}

\author{Edward Raff}
\affiliation{%
  \institution{University of Maryland Baltimore County \\
      Booz Allen Hamilton}
  % \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Baltimore}
  \country{USA}}
\email{Raff\_Edward@bah.com}

\author{Francis Ferraro}
\affiliation{%
  \institution{University of Maryland Baltimore County}
  % \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Baltimore}
  \country{USA}}
\email{ferraro@umbc.edu}

\author{Cynthia Matuszek}
\affiliation{%
  \institution{University of Maryland Baltimore County}
  \city{Baltimore}
  \country{USA}}
\email{cmat@umbc.edu}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.

\begin{abstract}
    Dealing with robots and other physical systems often leads to large amounts of complex, multimodal data, which may not be reliably available throughout the span of an interaction.
    Our study is motivated by needs in robotics, where an agent may have many sensors and thus many modalities by which a human may interact.
    % , such as voice, text, images, and depth.
    Tying these various modalities together relates to a task in grounded language understanding in which natural language is used as a retrieval query against objects in a physical environment.
    However, current work has neglected the real-world constraint that sensors fail due to obstructions or adverse conditions.
    We propose a generalized distance-based loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable. 
    We demonstrate the usability of our model on a grounded language object retrieval scenario. We leverage four modalities including vision, depth sensing, text, and speech, and demonstrate that our model learns faster than state-of-the-art baselines, while outperforming them when less training data is available.
    The code is publicly available on GitHub and will be included for the camera-ready version.
%%% END of previous abstract

    % TMLR abstract
    % Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities with which a human may interact, both to communicate a desired goal and for the agent to recognize a desired target object. For such problems, there has been little research on integrating more than two modalities. While there have been widely popular works based on cross-entropy, there is an entirely separate family of approaches based on explicit geometric alignment known as contrastive learning. However, to the best of our knowledge there has been no work on combining the two approaches for multimodal learning. We propose to combine both families of approaches and argue that the two are complementary. We introduce \textit{extended multimodal alignment (EMMA)}, a geometric (distance-based) method combined with a contrastive loss function. Our method can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, compounded by the challenge of retrieval when a modality becomes unavailable. We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command. We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model converges approximately 2.3 times faster than previous strong baselines, and out-performs or is strongly competitive with state-of-the-art contrastive learning. % (in 80\% less time).
    % END of TMLR abstract

\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Machine Learning~Computer Vision</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Machine Learning~Natural Language Processing</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Information Retrieval~Object Detection</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Deep Learning~Neural Networks</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

% \ccsdesc[500]{Machine Learning~Computer Vision}
\ccsdesc{Machine Learning~Computer Vision}
\ccsdesc{Machine Learning~Natural Language Processing}
\ccsdesc{Information Retrieval~Object Detection}
% \ccsdesc[100]{Deep Learning~Neural Networks}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{deep learning, neural networks, multimodal learning}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{comment}

\begin{teaserfigure}
  \includegraphics[width=\textwidth]{sampleteaser}
  \caption{Seattle Mariners at Spring Training, 2010.}
  \Description{Enjoying the baseball game from the third-base
  seats. Ichiro Suzuki preparing to bat.}
  \label{fig:teaser}
\end{teaserfigure}
\end{comment}

% \received{\todokdinline{Should we remove these?}20 February 2007}
% \received[revised]{12 March 2009}
% \received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

% \input{Inbox.tex}

\section{Introduction}
\label{intro}
Humans use multiple sensory inputs to understand, interact with, and retrieve information from the world around them. It is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources. The benefits of integrating multiple modalities are twofold: first, complementary information from different modalities can help with understanding the world, and second, additional modalities can help when one or more sources of data become unavailable.
% 
Grounded language understanding-based object retrieval, in which natural language is used as a query against objects in a physical environment, allows an intuitive mechanism by which users can instruct agents to engage in tasks. Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\citep{hong2021gilbert,Zhuge_2021_CVPR_VLP}.

Although physical agents such as robots typically have access to multiple sensory and interactive modalities, most approaches use at most two sensory inputs (e.g., RGB plus depth images) with single labels. Simultaneously using additional modalities and labels is an underexplored area.
% , in part due to the domain-specific nature of such \textit{n}-ary learning approaches. With the modern proliferation of audio and text based communication and home agents (e.g., Alexa/Google Home), 
% there is a growing need to handle more modalities and their potential failures. 
However, one difficulty of working with complex multimodal data is the increased likelihood that one or more modalities may have missing information. Hardware can become damaged or defective, sensors can get blocked or obstructed, and various adverse conditions can remove a modality from use. Current multimodal approaches are typically not robust to the loss of one or more modalities at test time.

In order to fully leverage multimodal training data, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities, while being robust to missing information. We consider the domain of grounded language-based object retrieval~\citep{hu2016natural, triplet_loss_2021_CVPR}, in which objects in an environment must be identified based on linguistic instructions. This can be considered a special case of image retrieval~\citep{huang2017deep, ma2020large, novak2015large, vo2019composing}. Previous approaches have explored combinations of sensors such as depth and RGB with labels provided by language~\citep{RichardsDarvishMatuszekCategoryFree20}. However, despite the multisensory nature of object retrieval, much of the existing work has not previously been extended to include an arbitrary number of modalities. The ultimate aim of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities.

In this work, we propose our new method called \textit{extended multimodal alignment (\ours{})}, a generalized geometric method combined with a cross-entropy based supervised contrastive loss function~\cite{NEURIPS2020_supervised_contrastive}, which we abbreviate \supcon{}. Our method can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data. We demonstrate steps towards this goal with a mechanism that learns quickly from a variety of input signals and is robust to ablation of inputs.
% This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs.

Our contributions are as follows:
% \begin{enumerate}

(1) We propose a new approach to multimodal learning by combining geometric and cross-entropy methods. Our method is designed to be applicable to additional modalities as they become available.

(2) Our method converges approximately 2.3 times faster than previous baselines, and out-performs state-of-the-art contrastive learning~\citep{chen2020simple} and supervised contrastive learning~\citep{NEURIPS2020_supervised_contrastive} when less training data is available.
    % and/or modalities are ablated.

(3) We demonstrate robustness in the face of missing input signals when one or more modalities are ablated at test time, and propose a simple averaging method to quantify performance for multimodal object retrieval scenarios.

(4) We demonstrate the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.

% \end{enumerate}

@@

There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case, for example in
% . In this task, language is used to formulate queries against datasets of images, 
% for example in text-and-image matching tasks for fashion data~\citep{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\citep{huang2017deep}, and 
general photographs of objects~\citep{ma2020large, novak2015large, hong2021gilbert}. Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\citep{vo2019composing}. While recent works have focused on grounding with respect to a knowledge set \citep{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning to perform grounding over an arbitrary number of input modalities.

The growing number of datasets that contain modalities beyond images and text demonstrates the importance and the challenge of this task. \citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition. 
% They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).
~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\citep{KebeAAAI2022}. 
\citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization. They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}. In this work, we focus on alignment, the process of identifying the direct relationship between elements across modalities.

We highlight a key difference with recent work like on multi-modal retrieval models~\cite{10.1145/3397271.3401232, 10.1145/3331184.3331213}. These retrieval models are class based, whereas our method is instance based, where different objects are not considered the same even if they have the same class. These works also do not consider performance in the case of failure of an input modality.

A number of different approaches have been proposed for linguistic-based alignment learning. \citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language. 
% Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video, and these embeddings are then mapped to a coarse space where text is also mapped. Their network respects specificity and comparability of different modalities. 
The loss function they define consists of two distance-based terms for dense and coarse spaces, differing from ours in that they do not handle arbitrarily many modalities and do not focus on robustness in the face of modality drop-outs. \citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks. \citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss \citep{Chechik:2010:LSO:1756006.1756042}, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.

There are two broad families of approaches to multimodal learning in the literature. One family is based on classification and cross-entropy losses~\cite{NEURIPS2020_supervised_contrastive, chen2020simple}, and the other is geometric~\cite{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}. In this paper we marry the two families and show that combining a geometric approach with cross-entropy based \supcon{} loss function is superior.

% For the classification part of our loss function we employ the supervised contrastive loss proposed by~\citet{NEURIPS2020_supervised_contrastive}. 
Our geometric loss function formulation is similar to lifted structured loss~\citep{songCVPR16LiftedStructured}. They take a geometric approach in a unimodal scenario and consider pairwise distances among all items in a batch, but we apply our idea to four modalities (and could incorporate more).
This extended geometric formulation results in a loss function that is similar to \citet{ICML22GeometricMultimodal}, where they 
% propose a geometric contrastive learning approach in which they 
fuse all embeddings as a central embedding and then minimize the distance between each embedding and that central embedding. However, their method does not take into account any notion of classification.

The geometric approach is also known as contrastive learning~\citep{qin2021world}, which is based on the idea of similarity and distance of elements of training data~\citep{Carvalho-cooking-triplet,triplet_loss_2021_CVPR,salvador2017cooking}. Triplet loss is also a special case of contrastive loss~\citep{NEURIPS2020_supervised_contrastive}. Contrastive loss is mostly used in self-supervised learning~\citep{bui2021self, alayrac2020self,chen2020simple}, with some exceptions such as \citet{NEURIPS2020_supervised_contrastive}. Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only). Our work is a specialization of contrastive loss that is focused on the multimodal input and language problem.

We note that standard triplet learning approaches often require an expensive \textit{mining} step to find harder negative samples to enable effective learning \citep{Hoffer2015,Schroff2015,DBLP:conf/eccv/2018-9,DBLP:conf/eccv/ZhaoJQLH18,Zhai2018}. This is problematic for throughput and scalability, as run time becomes quadratic in batch size. Our approach does not require mining to obtain good performance, alleviating these issues along with their complicated history in reproducibility and parameter tuning \citep{Musgrave2020,Raff2020c,Raff2019_quantify_repro}.
